---
title: "5_Solutions"
author: "Evan Muzzall"
date: "2023-05-12"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

# Challenge 1 - dataframes

1. Open a new script a create a dataframe that contains 6 rows and 3 columns by following the instructions above.

2. Advanced: what is the difference between a data frame and a tidyverse tibble? 

```{r}
# 1. 
gpa <- c(3.5, 4.0, 3.3, 2.9, 3.8, 3.5)
school <- c("UC Berkeley", "Stanford", "Stanford", "Stanford", "UC Berkeley", "UC Berkeley")
graduated <- c(T, F, T, F, F, T)

students <- data.frame(gpa, school, graduated)
students

# or directly within the call to data.frame()
students <- data.frame(gpa = c(3.5, 4.0, 3.3, 2.9, 3.8, 3.5), 
                       school = c("UCB", "SU", "SU", "SU", "UCB", "UCB"), 
                       graduated = c(T, F, T, F, F, T))
students

# 2. 
# https://jtr13.github.io/cc21fall1/tibble-vs.-dataframe.html#:~:text=10%209.,string%20into%20factor%20by%20default)
```

# Challenge 2 - subsetting

1. Open a new script a create a subset that contains data from DC or stops less than or equal to 7000 and just columns "city" and "traffic_stops"

2. Advanced: use the `filter()` and `select()` functions from the `dplyr` R package to do the same thing (hint: see the dplyr section below!)

```{r}
#1. 
# recreate the dataframe if necessary
police <- data.frame(traffic_stops = c(8814, 9915, 9829, 10161, 6810, 8991), 
                     city = as.factor(c("SF", "DC", "DC", "DC", "SF", "SF")),
                     year = as.integer(c(2000, 2000, 2001, 2002, 2001, 2002)))
police

# the base R way
dc_7000 <- police[police$city == "DC" |
                    police$traffic_stops <= 7000, c("city", "traffic_stops")]
dc_7000

# or, the dplyr way!
library(dplyr)
dc_7000 <- police %>%
  filter(city == "DC" | traffic_stops <= 7000) %>%
  select(city, traffic_stops)
dc_7000
```

# Challenge 3 - `group_by()` and `mutate()`

While `summarize()` will create a new tibble, `mutate()` creates a new column(s) in an existing tibble. 

Open a new script and use the help file for the `mutate()` function that, combined with `group_by()`, creates a new column that computes the average stops per year for the `police` dataframe. (hint: assume there are 365 days in a year!)

```{r}
# 1.
# recreate the dataframe if necessary
police <- data.frame(traffic_stops = c(8814, 9915, 9829, 10161, 6810, 8991), 
                     city = as.factor(c("SF", "DC", "DC", "DC", "SF", "SF")),
                     year = as.integer(c(2000, 2000, 2001, 2002, 2001, 2002)))

avg_stops <- police %>% 
  group_by(year) %>% 
  mutate(avg = mean(traffic_stops))
avg_stops
```

# Challenge 4 - .txt import

1. Import the "frankenstein.txt" file from the "novels" subfolder. Save it in a variable named `frank`

```{r}
# 1. 
# Read .txt file to a string
frank <- paste(readLines("data/raw/novels/frankenstein.txt"), collapse="\n")

# With the readr package
# install.packages("readr")
frank <- readr::read_file("data/raw/novels/frankenstein.txt")

# Read the .txt file as a character string
txt_file <- "data/raw/novels/frankenstein.txt"
frank <- readChar(txt_file, file.info(txt_file)$size)
```

# Challenge 5 - line plot

1. Use `geom_line()` to create a line plot of mean lifeExpectancy through time FOR EACH CONTINENT using the gap dataset loaded below. Before plotting however, use dplyr to pipe in your `group_by()` and `summarize()` functions directly into ggplot. 

```{r}
library(dplyr)
library(ggplot2)
gap <- read.csv("data/raw/gapminder-FiveYearData.csv")

gap %>% group_by(continent, year) %>%
  summarize(mean_LE = mean(lifeExp, na.rm = TRUE)) %>% 
  ggplot() + geom_line(mapping = aes(x = year, y = mean_LE, 
                                     color = continent), 
                       lwd = 2) + xlab("Year") + ylab("Mean life expectancy") +
  theme_minimal() 
```

# Challenge 6 - summary statistics

1. Install and library the "psych" package and call the help file `?describeBy` to see how to produce this function's default summary statistics as grouped by continent.
```{r}
# install.packages("psych")
library(psych)

sum_stats <- psych::describeBy(gap, group = gap$continent)
sum_stats
```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```